Comparing Hydrogen and Jet-A for an N+3 Turbofan with Water Recirculation using Gradient-Based Optimization

### 2023-05-10 conference-paper/main.tex
More changes

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{siunitx}
\sisetup{group-separator = {,}}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{float}
\usepackage{amsmath}
\usepackage[version=4]{mhchem}
\usepackage{longtable,tabularx}
\usepackage{placeins}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{latexsym}
\usepackage{multicol}
\usepackage{subcaption}
\usepackage[sort&compress,numbers]{natbib} % For bibtex \citet, \citep
\usepackage{hypernat} % To get natbib to play nicely with hyperref
\usepackage{doi} % For getting hyperlinked DOI in the references
\usepackage{hyperref}

@@

% Message 2: Background and references to support using H2 and water injection in HBTF engines
% [x] TODO: AL-PA, Need to find a journal/conference paper to cite for the following statement.  Maybe find one or more papers other than just the NASA report?
Water injection is the process of introducing water upstream of the combustor as finely atomized droplets.
NASA, Boeing, and Rolls-Royce studied this concept and suggested that this technique reduces the NOx emissions as much as 47 percent~\cite{Daggett2010}.
Additionally, water injection improves fuel efficiency and thrust output with lower combustion temperatures that can improve the lifetime of turbine blades and reduce noise~\cite{Daggett2010}.
% [x] TODO: AL-PA, Should try to find a citation for this if one exists.
Traditional propulsion systems that burn hydrocarbon fuels would require external demineralized water storage on the aircraft for water injection~\cite{Mourouzidis2015}.
The added weight of tanks, pumping, and ducting makes this concept infeasible for a conventional aircraft.
The main product of hydrogen combustion is water vapor~\cite{Strom2002}.
We can recover the vapor from the exhaust stream with condensers and recirculate it upstream to eliminate additional storage requirements.
If hydrogen is used as the fuel then it can be used as a heat sink for the condenser~\cite{Boggia2002}.
% Condensing water vapor from the exhaust stream of hydrogen combustion and recirculating it eliminates any additional storage requirements.
This system represents a closed loop water feedback system inside the propulsion cycle.
% This allows for the theoretical design of a closed loop water feedback system inside the propulsion cycle.
Pratt and Whitney are actively researching this technology to increase the efficiency of hydrogen-powered propulsion~\cite{arpa-e_2021}.
% They suggest that water vapor can be recovered by utilizing the thermal properties of hydrogen to condense some of the water in the turbofan exhaust.

@@

% NOTE: AL-PA, I turned this into a subfigure to save some space
\begin{figure}[hbt!]
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{injector.pdf}
    \caption{
      Water from the extractor is simply injected into the core flow upstream of the high-pressure compressor.
    }
    \label{fig:injector}
  \end{subfigure}
  \hspace{2pt}
  \begin{subfigure}[t]{0.49\textwidth}
    \includegraphics[width=\textwidth]{extractor.pdf}
    \caption{
      We extract a fraction of the mass flow of water determined by the molar composition.
      The extracted water is routed from duct 5 back upstream to the injector and the rest is exhausted out the nozzle.
    }
    \label{fig:extractor}
  \end{subfigure}
  % [x] TODO: AL-PA, Might want to add a bit more to this caption if you think it's necessary.
  \caption{Injector (left) and extractor(right) diagrams that show the inputs and output streams of each component.
    The inputs and outputs show which flow values are used to compute the amount of recirculated water.}
  \label{fig:extract_inject}
\end{figure}

@@

\begin{equation}
  \mathrm{TSEC} = \frac{\dot{m}_{\mathrm{fuel}} \mathrm{LHV}}{F_{\mathrm{thrust}}} = \mathrm{TSFC} \times \mathrm{LHV}
  \label{eq:tsec}
\end{equation}
% In this work we are concerned with minimizing fuel burn of the engine subject to thrust constraints.
% [x] TODO: AL-PA, TSEC is not the objective.  Fuel burn is the objective.  TSEC is the efficiency metric of interest.
TSEC will be used as a efficiency metric at the cruise condition to compare the two fuels in the results.
% [x] TODO: AL-PA, explain the important aspects of the XDSM in writing before referencing the figure.
The fuel properties used in the engine model are given in~\ref{fuel_props}.

% [x] TODO: AL-PA, Why are some of the fist words in the descriptions capitalized but not others?  I think this should be consistent.
% [x] TODO: AL-PA, Change the lower heating value references to footnotes. i.e \footnote{\url{...}}
\begin{table}[hbt!]
  \centering
  \caption{
    N+3 model fuel properties.
    The lower heating values are given for each fuel used to compute TSEC.}
  \begin{tabular}{r r r l}
    \hline
    Parameter              & Value & Units        & Description                                                                                                            \\
    \hline
    $\rm{LHV}_{\rm{JetA}}$ & 18564 & \si{BTU/lbm} & Lower heating value of Jet-A~\footnote{\url{https://www.smartcockpit.com/docs/Jet_Fuel_Characteristics.pdf}}           \\
    $\rm{LHV}_{\rm{H2}}$   & 51591 & \si{BTU/lbm} & Lower heating value of H2~\footnote{\url{https://www.engineeringtoolbox.com/fuels-higher-calorific-values-d_169.html}} \\
    \hline
  \end{tabular}
  \label{fuel_props}
\end{table}

@@

@@

\begin{equation}
  V_{\rm{ratio}} = \frac{V_{\rm{core,ideal}}C_{v,\rm{core}}}{V_{\rm{bypass,ideal}}C_{v,\rm{bypass}}}
  \label{eq:vratio}
\end{equation}
where $V_\text{ideal}$ is the ideal nozzle exit velocity and $C_v$ is the nozzle velocity coefficient that accounts for non-ideal nozzle effects.
The final design variable is the water recovery fraction at CRZ ($X_{\rm{H_2O,CRZ}}$).
% [x] TODO: AL-PA, optimizing for fuel burn subject to net thrust constraint...
% [x] TODO: AL-PA, this isn't the only reason we consider just the water recovery at cruise.  It's also the longest flight segment and contributes the most to long-term performance.
We only include the water recovery fraction at cruise because it has the largest impact on the fuel mass flow rate objective function.
Additionally, the CRZ operating condition is the longest flight segment and contributes the most to long-term engine performance.
The range for each design variable and the associated units are shown in Table~\ref{tab:dv_table}.
% We add the water recovery fraction only at the CRZ point as a design variable since we are optimizing $\dot{m}_{\rm{fuel,CRZ}}$.

% [x] TODO: AL-PA, You already explained the flight conditions in the multipoint section so not necessary here
% [x] TODO: AL-PA, I would describe the constraints in more detail here.  This might be a good spot to talk about the constraints imposed by the balance comps at the top level.
%                  - i.e the SLS and CRZ net thrust constraints
%                  - i.e how the BPR is set by the nozzle velocity ratio
%                  - i.e The interdependence of the TOC mass flow rate and the RTO temperature
As aforementioned in Section~\ref{sec:method}, the N+3 model uses balance relationships to enforce design, conservation, and performance requirements.
To create a reduced space optimization problem, the N+3 model imposes a set of constraints using additional balance relationships.
These constraints are satisfied using a Newton solver at the analysis level instead of the optimization level.
The advantage of the reduced space approach is that the design is feasible at the completion of each optimization iteration.
\citet{Martins2013} classify this architecture as multidisciplinary design feasible (MDF).
The balance relationships constrain the net thrust at the SLS and CRZ conditions, the BPR at TOC, and the turbine inlet temperature ($T_4$) at RTO.
We summarize the constraints as residuals (r) and the associated implicit states ($\vec{u}$) in Equation~\eqref{eq:balance_cons}.
\begin{equation}
  \label{eq:balance_cons}
  r(\vec{u}) \rightarrow
  \begin{cases}
    \setlength{\tabcolsep}{1pt}
    \begin{tabular}{l l}
      $r(\text{BPR}_\text{TOC})$    & $= (V_\text{ratio,TOC})_\text{des} - V_\text{ratio,TOC} = 0$, \\
      $r(\dot{m}_\text{total,TOC})$ & $= (T_{4,\text{RTO}})_\text{des} - T_{4,\text{RTO}} = 0$,     \\
      $r(\text{FAR}_\text{CRZ})$    & $= 0.9 F_\text{net, TOC} - F_\text{net, CRZ} = 0$,            \\
      $r(\text{FAR}_\text{SLS})$    & $= 1.2553 F_\text{net,RTO} - F_\text{net, SLS} = 0$.          \\
    \end{tabular}
  \end{cases}
\end{equation}

@@

% [x] TODO: AL-PA, Define the T4 ratio with an equation and reference the equation in the table
% [x] TODO: AL-PA, I thought we were only doing recovery fractions at CRZ??
% [x] TODO: AL-PA, Make sure you reference the equations for the DVs/cons in the descriptions so the reader can connect them.
% [x] TODO: PA-PA, explain variable ranges
\begin{table}[hbt!]
  \centering
  \caption{
    Multipoint optimization problem definition.
    The objective function is the fuel flow rate at the cruise condition.
    The constraints are net thrust and engine diameter constraint at the design point, TOC.
  }
  \small
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{r l l r l}
    \toprule
                    & Variable/Function              & Description                                                              & Units          & Quantity \\
    \hline
    minimize        & $\dot{m}_{\rm{fuel,CRZ}} $     & Fuel flow rate at CRZ                                                    & \si{lbm/s}     & 1        \\
                    &                                &                                                                          &                &          \\
    with respect to & $X_{\rm{H_2O,CRZ}}$            & Water recovery mass fraction at CRZ                                      & -              & 1        \\
                    & $\rm{PR}_{\rm{fan,TOC}}$       & TOC fan pressure ratio                                                   & -              & 1        \\
                    & $\rm{PR}_{\rm{LPC,TOC}}$       & TOC low-pressure compressor pressure ratio                               & -              & 1        \\
                    & $\rm{OPR}_{\rm{TOC}}$          & TOC overall pressure ratio                                               & -              & 1        \\
                    & $T_{4,\rm{RTO}}$               & RTO combustor temperature                                                & \si{\degree R} & 1        \\
                    & $T_{4,\rm{ratio}}$             & TOC-to-RTO temperature ratio (Equation~\eqref{t4ratio})                  & -              & 1        \\
                    & $V_{\rm{ratio,CRZ}}$           & Core-to-bypass nozzle velocity ratio at CRZ (Equation~\eqref{eq:vratio}) & -              & 1        \\
    \cline{3-5}
                    &                                & Total                                                                    &                & 7        \\
                    &                                &                                                                          &                &          \\
    subject to      & $F_{\rm{net,TOC}} \geq 5800.0$ & Target net thrust at TOC                                                 & \si{lbf}       & 1        \\
                    & $D_{\rm{fan}} \leq 100$        & Maximum Fan Diameter                                                     & \si{inch^2}    & 1        \\
    \cline{3-5}
                    &                                & Total                                                                    &                & 2        \\
    \bottomrule
  \end{tabular}
  \label{tab:opt_problem}
\end{table}

\begin{table}[hbt!]
  \centering
  \caption{Design variable ranges for the optimization problem.
  }
  \small
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{l l l l}
    Design Variable          & Lower  & Upper  & Units     \\
    \toprule
    $X_{\rm{H_2O,CRZ}}$      & 0.0    & 1.0    & -         \\
    $\rm{PR}_{\rm{fan,TOC}}$ & 1.2    & 1.4    & -         \\
    $\rm{PR}_{\rm{LPC,TOC}}$ & 2.5    & 4.0    & -         \\
    $\rm{OPR}_{\rm{TOC}}$    & 40.0   & 70.0   & -         \\
    $T_{4,\rm{ratio}}$       & 0.5    & 0.95   & -         \\
    $T_{4,RTO}$              & 3000.0 & 3600.0 & $^\circ$R \\
    $V_{ratio,CRZ}$          & 1.35   & 1.45   & -         \\
    \bottomrule
  \end{tabular}
  \label{tab:dv_table}
\end{table}

% Blurb about PyOptSparse and SNOPT
% NOTE: AL-PA, This can all be part of the optimization problem section you don't need a new subsection
% \subsection{Optimization Software}
% NOTE: AL-PA, "pyOptSparse" not "PyOptSparse", i.e the first "p" is not capitalized.
% [x] TODO: AL-AL, need to cite pyOptSparse, SNOPT, and OpenMDAO
% NOTE: AL-PA, I re-wrote most of this to be more concise.
The N+3 pyCycle model is implemented in OpenMDAO~\cite{Gray2019a} to perform multidisciplinary gradient-based optimization with analytic coupled derivatives.
We use pyOptSparse~\cite{Wu2020a} to facilitate the use of state-of-the-art optimization software through a unified python interface.
We solve the optimization problem listed in Table~\ref{tab:opt_problem} with SNOPT~\cite{Gill2005a}, a gradient-based sequential quadratic programming (SQP) algorithm for large-scale constrained problems.

% NOTE: AL-PA, this was confusing and I don't think it add much useful information that the readers will actually understand.

% During initial optimization studies of this model, it was found that when $T_{4,\rm{RTO}}$ and $X_{\rm{H_2O,CRZ}}$ are design variables at the same time they had opposite effects on the design.
% The optimizer increases $X_{\rm{H_2O,CRZ}}$ which allows for a lower optimal $T_{4,\rm{RTO}}$.
% This is likely due to the interdependency between engine mass flow rate and $T_{4,\rm{RTO}}$.
% As the recovery water fraction increases less mass flow rate is required through the engine which lowers the required $T_{4,\rm{RTO}}$ for a required level of thrust.
% It was also observed that $X_{\rm{H_2O,CRZ}}$ is increased until the model breaks due to the CEA solver not being able resolve the increased \ce{H2O} molecules that are injected.
% Since $T_{4,\rm{RTO}}$ would be pushed down to its lower limit, it was set as the optimal value from the optimizations with no water recovery.
% This allowed the optimizer to find the maximum $X_{\rm{H_2O,CRZ}}$ that would allow the most water to be recirculated.
% The upper bound on $X_{\rm{H_2O,CRZ}}$ was found to be around 30\% for Jet-A and 17\% for \ce{H2}.

\section{Results}
\label{sec:results}

@@

\begin{table}[hbt!]
  \centering
  \caption{Optimization results of the N+3 engine for four optimizations using Jet-A and hydrogen fuel with and without water recirculation.
  }
  \small
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{r l l l l l l}
                & Variable/Function         & Jet-A Dry & Jet-A Wet & \ce{H2} Dry & \ce{H2} Wet & Units      \\
    \toprule
    objective   & $\dot{m}_{\rm{fuel,CRZ}}$ & 0.639     & 0.606     & 0.233       & 0.221       & \si{lbm/s} \\
    % & $\rm{TSEC}_{\rm{CRZ}}$    & 8186.254  & 7766.338  & 8295.236    & 7859.635    & \si{lbm/hr/lbf} \\
    \hline
    variables   & $x_{H2O,CRZ}$             & -         & 0.3       & -           & 0.17        & -          \\
                & $\rm{PR}_{fan,TOC}$       & 1.285     & 1.292     & 1.288       & 1.297       & -          \\
                & $\rm{PR}_{LPC,TOC}$       & 4.0       & 4.0       & 4.0         & 4.0         & -          \\
                & $\rm{OPR}_{\rm{TOC}}$     & 61.012    & 61.188    & 60.660      & 61.952      & -          \\
                & $T_{4,ratio}$             & 0.913     & 0.911     & 0.918       & 0.915       & -          \\
                & $T_{4,RTO}$               & 3220.574  & -         & 3204.071    & -           & $^\circ$R  \\
                & $V_{ratio,CRZ}$           & 1.35      & 1.35      & 1.35        & 1.35        & -          \\
    \hline
    constraints & $F_{net,TOC}$             & 5800.0    & 5800.0    & 5800.0      & 5800.0      & lbf        \\
                & $D_{Fan}$                 & 100.0     & 100.0     & 100.0       & 100.0       & $in^2$     \\
    \bottomrule
  \end{tabular}
  \label{tab:dv_opt}
\end{table}

@@

We varied the total pressure loss across the duct to simulate drag from an increasingly large condenser.

>>>>

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{siunitx}
\sisetup{group-separator = {,}}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{float}
\usepackage{amsmath}
\usepackage[version=4]{mhchem}
\usepackage{longtable,tabularx}
\usepackage{placeins}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{latexsym}
\usepackage{multicol}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage[sort&compress,numbers]{natbib} % For bibtex \citet, \citep
\usepackage{hypernat} % To get natbib to play nicely with hyperref
\usepackage{doi} % For getting hyperlinked DOI in the references

@@

% Message 2: Background and references to support using H2 and water injection in HBTF engines
% [x] TODO: AL-PA, Need to find a journal/conference paper to cite for the following statement.  Maybe find one or more papers other than just the NASA report?
Water injection is the process of introducing water upstream of the combustor as finely atomized droplets.
NASA, Boeing, and Rolls-Royce studied this concept and suggested that this technique reduces the NOx emissions as much as 47 percent~\cite{Daggett2010}.
Additionally, water injection improves fuel efficiency and thrust output with lower combustion temperatures that can improve the lifetime of turbine blades and reduce noise~\cite{Daggett2010}.
% [x] TODO: AL-PA, Should try to find a citation for this if one exists.
Traditional propulsion systems that burn hydrocarbon fuels would require external demineralized water storage on the aircraft for water injection~\cite{Mourouzidis2015}.
The added weight of tanks, pumping, and ducting makes this concept infeasible for a conventional aircraft.
The main product of hydrogen combustion is water vapor~\cite{Strom2002}.
We can recover the vapor from the exhaust stream with condensers and recirculate it upstream to eliminate additional storage requirements.
If hydrogen is used as the fuel then it can be used as a heat sink for the condenser~\cite{Boggia2002}.
% Condensing water vapor from the exhaust stream of hydrogen combustion and recirculating it eliminates any additional storage requirements.
This system represents a closed water feedback loop inside the propulsion cycle.
% This allows for the theoretical design of a closed loop water feedback system inside the propulsion cycle.
Pratt and Whitney are actively researching this technology to increase the efficiency of hydrogen-powered propulsion~\cite{arpa-e_2021}.
% They suggest that water vapor can be recovered by utilizing the thermal properties of hydrogen to condense some of the water in the turbofan exhaust.

@@

% NOTE: AL-PA, I turned this into a subfigure to save some space
\begin{figure}[hbt!]
    \centering
    \begin{subfigure}[t]{0.49\textwidth}
        \includegraphics[width=\textwidth]{injector.pdf}
        \caption{
            Water from the extractor is simply injected into the core flow upstream of the high-pressure compressor.
        }
        \label{fig:injector}
    \end{subfigure}
    \hspace{2pt}
    \begin{subfigure}[t]{0.49\textwidth}
        \includegraphics[width=\textwidth]{extractor.pdf}
        \caption{
            We extract a fraction of the mass flow of water determined by the molar composition.
            The extracted water is routed from Duct5 back upstream to the injector and the rest is exhausted out the nozzle.
        }
        \label{fig:extractor}
    \end{subfigure}
    % [x] TODO: AL-PA, Might want to add a bit more to this caption if you think it's necessary.
    \caption{Injector (left) and extractor(right) diagrams that show the inputs and output streams of each component.
        The inputs and outputs show which flow values are used to compute the amount of recirculated water.}
    \label{fig:extract_inject}
\end{figure}

@@

\begin{equation}
    \mathrm{TSEC} = \frac{\dot{m}_{\mathrm{fuel}} \mathrm{LHV}}{F_{\mathrm{thrust}}} = \mathrm{TSFC} \times \mathrm{LHV}
    \label{eq:tsec}
\end{equation}
% In this work we are concerned with minimizing fuel burn of the engine subject to thrust constraints.
% [x] TODO: AL-PA, TSEC is not the objective.  Fuel burn is the objective.  TSEC is the efficiency metric of interest.
TSEC will be used as a efficiency metric at the cruise condition to compare the two fuels in the results.
% [x] TODO: AL-PA, explain the important aspects of the XDSM in writing before referencing the figure.
The fuel properties used in the engine model are given in~\ref{fuel_props}~\footnote{\url{https://www.smartcockpit.com/docs/Jet_Fuel_Characteristics.pdf}}~\footnote{\url{https://www.engineeringtoolbox.com/fuels-higher-calorific-values-d_169.html}}.

% [x] TODO: AL-PA, Why are some of the fist words in the descriptions capitalized but not others?  I think this should be consistent.
% [x] TODO: AL-PA, Change the lower heating value references to footnotes. i.e \footnote{\url{...}}
\begin{table}[hbt!]
    \centering
    \caption{
        N+3 model fuel properties.
        The lower heating values are given for each fuel used to compute TSEC.}
    \begin{tabular}{r r r l}
        \hline
        Parameter              & Value & Units        & Description                  \\
        \hline
        $\rm{LHV}_{\rm{JetA}}$ & 18564 & \si{BTU/lbm} & Lower heating value of Jet-A \\
        $\rm{LHV}_{\rm{H2}}$   & 51591 & \si{BTU/lbm} & Lower heating value of H2    \\
        \hline
    \end{tabular}
    \label{fuel_props}
\end{table}

@@

\begin{table}[hbt!]
    \centering
    \caption{
        Multipoint optimization problem definition.
        The objective function is the fuel flow rate at the cruise condition.
        The constraints are net thrust and engine diameter constraint at the design point, TOC.
    }
    \small
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{r l l r l}
        \toprule
                        & Variable/Function              & Description                                                              & Units            & Quantity \\
        \hline
        minimize        & $\dot{m}_{\rm{fuel,CRZ}} $     & Fuel flow rate at CRZ                                                    & \unit{lbm/s}     & 1        \\
                        &                                &                                                                          &                  &          \\
        with respect to & $X_{\rm{H_2O,CRZ}}$            & Water recovery mass fraction at CRZ                                      & -                & 1        \\
                        & $\rm{PR}_{\rm{fan,TOC}}$       & TOC fan pressure ratio                                                   & -                & 1        \\
                        & $\rm{PR}_{\rm{LPC,TOC}}$       & TOC low-pressure compressor pressure ratio                               & -                & 1        \\
                        & $\rm{OPR}_{\rm{TOC}}$          & TOC overall pressure ratio                                               & -                & 1        \\
                        & $T_{4,\rm{RTO}}$               & RTO combustor temperature                                                & \unit{\degree R} & 1        \\
                        & $T_{4,\rm{ratio}}$             & TOC-to-RTO temperature ratio (Equation~\eqref{t4ratio})                  & -                & 1        \\
                        & $V_{\rm{ratio,CRZ}}$           & Core-to-bypass nozzle velocity ratio at CRZ (Equation~\eqref{eq:vratio}) & -                & 1        \\
        \cline{3-5}
                        &                                & Total                                                                    &                  & 7        \\
                        &                                &                                                                          &                  &          \\
        subject to      & $F_{\rm{net,TOC}} \geq 5800.0$ & Target net thrust at TOC                                                 & \unit{lbf}       & 1        \\
                        & $D_{\rm{fan}} \leq 100$        & Maximum Fan Diameter                                                     & \unit{inch^2}    & 1        \\
        \cline{3-5}
                        &                                & Total                                                                    &                  & 2        \\
        \bottomrule
    \end{tabular}
    \label{tab:opt_problem}
\end{table}

@@

\begin{equation}
    V_{\rm{ratio}} = \frac{V_{\rm{core,ideal}}C_{v,\rm{core}}}{V_{\rm{bypass,ideal}}C_{v,\rm{bypass}}}
    \label{eq:vratio}
\end{equation}
where $V_\text{ideal}$ is the ideal nozzle exit velocity and $C_v$ is the nozzle velocity coefficient that accounts for non-ideal nozzle effects.
The final design variable is the water recovery fraction at CRZ ($X_{\rm{H_2O,CRZ}}$).
% [x] TODO: AL-PA, optimizing for fuel burn subject to net thrust constraint...
% [x] TODO: AL-PA, this isn't the only reason we consider just the water recovery at cruise.  It's also the longest flight segment and contributes the most to long-term performance.
We only include the water recovery fraction at cruise because it has the largest impact on the fuel mass flow rate objective function.
Additionally, the CRZ operating condition is the longest flight segment and contributes the most to long-term engine performance.
The range for each design variable and the associated units are shown in Table~\ref{tab:dv_table}.
% We add the water recovery fraction only at the CRZ point as a design variable since we are optimizing $\dot{m}_{\rm{fuel,CRZ}}$.
\begin{table}[hbt!]
    \caption{Design variable ranges for the optimization problem.
    \small
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{l l l l}
        Design Variable          & Lower  & Upper  & Units            \\
        \toprule
        $X_{\rm{H_2O,CRZ}}$      & 0.0    & 1.0    & -                \\
        $\rm{PR}_{\rm{fan,TOC}}$ & 1.2    & 1.4    & -                \\
        $\rm{PR}_{\rm{LPC,TOC}}$ & 2.5    & 4.0    & -                \\
        $\rm{OPR}_{\rm{TOC}}$    & 40.0   & 70.0   & -                \\
        $T_{4,\rm{ratio}}$       & 0.5    & 0.95   & -                \\
        $T_{4,RTO}$              & 3000.0 & 3600.0 & \unit{\degree R} \\
        $V_{ratio,CRZ}$          & 1.35   & 1.45   & -                \\
        \bottomrule
    \end{tabular}
    \label{tab:dv_table}
\end{table}
% [x] TODO: AL-PA, You already explained the flight conditions in the multipoint section so not necessary here
% [x] TODO: AL-PA, I would describe the constraints in more detail here.  This might be a good spot to talk about the constraints imposed by the balance comps at the top level.
%                  - i.e the SLS and CRZ net thrust constraints
%                  - i.e how the BPR is set by the nozzle velocity ratio
%                  - i.e The interdependence of the TOC mass flow rate and the RTO temperature
As aforementioned in Section~\ref{sec:method}, the N+3 model uses balance relationships to enforce design, conservation, and performance requirements.
To create a reduced space optimization problem, the N+3 model imposes a set of constraints using additional balance relationships.
These constraints are satisfied using a Newton solver at the analysis level instead of the optimization level.
The advantage of the reduced space approach is that the design is feasible at the completion of each optimization iteration.
\citet{Martins2013} classify this architecture as multidisciplinary design feasible (MDF).
The balance relationships constrain the net thrust at the SLS and CRZ conditions, the BPR at TOC, and the turbine inlet temperature ($T_4$) at RTO.
We summarize the constraints as residuals (r) and the associated implicit states ($\vec{u}$) in Equation~\eqref{eq:balance_cons}.
\begin{equation}
    \label{eq:balance_cons}
    r(\vec{u}) \rightarrow
    \begin{cases}
        \setlength{\tabcolsep}{1pt}
        \begin{tabular}{l l}
            $r(\text{BPR}_\text{TOC})$    & $= (V_\text{ratio,TOC})_\text{des} - V_\text{ratio,TOC} = 0$, \\
            $r(\dot{m}_\text{total,TOC})$ & $= (T_{4,\text{RTO}})_\text{des} - T_{4,\text{RTO}} = 0$,     \\
            $r(\text{FAR}_\text{CRZ})$    & $= 0.9 F_\text{net, TOC} - F_\text{net, CRZ} = 0$,            \\
            $r(\text{FAR}_\text{SLS})$    & $= 1.2553 F_\text{net,RTO} - F_\text{net, SLS} = 0$.          \\
        \end{tabular}
    \end{cases}
\end{equation}

@@

% [x] TODO: AL-PA, Define the T4 ratio with an equation and reference the equation in the table
% [x] TODO: AL-PA, I thought we were only doing recovery fractions at CRZ??
% [x] TODO: AL-PA, Make sure you reference the equations for the DVs/cons in the descriptions so the reader can connect them.
% [x] TODO: PA-PA, explain variable ranges

% Blurb about PyOptSparse and SNOPT
% NOTE: AL-PA, This can all be part of the optimization problem section you don't need a new subsection
% \subsection{Optimization Software}
% NOTE: AL-PA, "pyOptSparse" not "PyOptSparse", i.e the first "p" is not capitalized.
% [x] TODO: AL-AL, need to cite pyOptSparse, SNOPT, and OpenMDAO
% NOTE: AL-PA, I re-wrote most of this to be more concise.
The N+3 pyCycle model is implemented in OpenMDAO~\cite{Gray2019a} to perform multidisciplinary gradient-based optimization with analytic coupled derivatives.
We use pyOptSparse~\cite{Wu2020a} to facilitate the use of state-of-the-art optimization software through a unified python interface.
We solve the optimization problem listed in Table~\ref{tab:opt_problem} with SNOPT~\cite{Gill2005a}, a gradient-based sequential quadratic programming (SQP) algorithm for large-scale constrained problems.
% NOTE: AL-PA, this was confusing and I don't think it add much useful information that the readers will actually understand.

% During initial optimization studies of this model, it was found that when $T_{4,\rm{RTO}}$ and $X_{\rm{H_2O,CRZ}}$ are design variables at the same time they had opposite effects on the design.
% The optimizer increases $X_{\rm{H_2O,CRZ}}$ which allows for a lower optimal $T_{4,\rm{RTO}}$.
% This is likely due to the interdependency between engine mass flow rate and $T_{4,\rm{RTO}}$.
% As the recovery water fraction increases less mass flow rate is required through the engine which lowers the required $T_{4,\rm{RTO}}$ for a required level of thrust.
% It was also observed that $X_{\rm{H_2O,CRZ}}$ is increased until the model breaks due to the CEA solver not being able resolve the increased \ce{H2O} molecules that are injected.
% Since $T_{4,\rm{RTO}}$ would be pushed down to its lower limit, it was set as the optimal value from the optimizations with no water recovery.
% This allowed the optimizer to find the maximum $X_{\rm{H_2O,CRZ}}$ that would allow the most water to be recirculated.
% The upper bound on $X_{\rm{H_2O,CRZ}}$ was found to be around 30\% for Jet-A and 17\% for \ce{H2}.
\section{Results}
\label{sec:results}

@@

\begin{table}[hbt!]
    \centering
    \caption{Optimization results of the N+3 engine for four optimizations using Jet-A and hydrogen fuel with and without water recirculation.
    }
    \small
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{r l l l l l l}
                    & Variable/Function         & Jet-A Dry & Jet-A Wet & \ce{H2} Dry & \ce{H2} Wet & Units        \\
        \toprule
        objective   & $\dot{m}_{\rm{fuel,CRZ}}$ & 0.639     & 0.606     & 0.233       & 0.221       & \unit{lbm/s} \\
        % & $\rm{TSEC}_{\rm{CRZ}}$    & 8186.254  & 7766.338  & 8295.236    & 7859.635    & \si{lbm/hr/lbf} \\
        \hline
        variables   & $x_{H2O,CRZ}$             & -         & 0.3       & -           & 0.17        & -            \\
                    & $\rm{PR}_{fan,TOC}$       & 1.285     & 1.292     & 1.288       & 1.297       & -            \\
                    & $\rm{PR}_{LPC,TOC}$       & 4.0       & 4.0       & 4.0         & 4.0         & -            \\
                    & $\rm{OPR}_{\rm{TOC}}$     & 61.012    & 61.188    & 60.660      & 61.952      & -            \\
                    & $T_{4,ratio}$             & 0.913     & 0.911     & 0.918       & 0.915       & -            \\
                    & $T_{4,RTO}$               & 3220.574  & -         & 3204.071    & -           & $^\circ$R    \\
                    & $V_{ratio,CRZ}$           & 1.35      & 1.35      & 1.35        & 1.35        & -            \\
        \hline
        constraints & $F_{net,TOC}$             & 5800.0    & 5800.0    & 5800.0      & 5800.0      & lbf          \\
                    & $D_{Fan}$                 & 100.0     & 100.0     & 100.0       & 100.0       & $in^2$       \\
        \bottomrule
    \end{tabular}
    \label{tab:dv_opt}
\end{table}

@@

We varied the total pressure loss across the duct to simulate an increasingly larger condenser jutting out into the duct.
